---
title: "Exploring ToothGrowth Dataset"
output: pdf_document
author: "Naren Vijayakumar"
fontsize: 11pt
geometry: margin=1in
---

```{r warning=FALSE, message=FALSE}
library(datasets)
library(dplyr)
```

Loading the required dataset
```{r}
mydf <- ToothGrowth
mydf$dose <- as.factor(mydf$dose)
```

This dataset is based on the study of effects of Vitamin C on tooth growth in guinea pigs. 
```{r}
names(mydf)
```
It conists of 3 fields. ***len*** corresponds to length of odontoblasts (teeth) in each of 10 guinea pigs, ***dose*** corresponds to dose levels and ***supp*** corresponds to delivery methods orange juice (OJ) or ascrobic acid (VC). Box plot for different dosage levels.

```{r}
table(mydf$supp, mydf$dose)
```

```{r,fig.align='center',fig.height=3,fig.width=7,message=FALSE,warning=FALSE}
ggplot(mydf, aes(x=supp,y=len)) +  geom_boxplot(aes(group=supp,fill=supp)) + 
facet_grid(. ~ dose) + labs(x="Supplement Type", y="Length of Teeth",title="Boxplot for Different Dosage Levels")
```

Let us perform t test for different dosage levels. Our Null Hypothesis **H0: there is no difference between supplement types** and alternate hypothesis **Ha: there is difference between supplement types**. We set ***alpha=0.05***. 

```{r}
# OJ at different dosage levels
OJ_05 <- mydf %>% filter(dose==0.5 & supp=="OJ") %>% select(len)
OJ_1 <- mydf %>% filter(dose==1.0 & supp=="OJ") %>% select(len)
OJ_2 <- mydf %>% filter(dose==2.0 & supp=="OJ") %>% select(len)
# VC at different dosage levels
VC_05 <- mydf %>% filter(dose==0.5 & supp=="VC") %>% select(len)
VC_1 <- mydf %>% filter(dose==1.0 & supp=="VC") %>% select(len)
VC_2 <- mydf %>% filter(dose==2.0 & supp=="VC") %>% select(len)
```


```{r}
# p-value at 0.5 dosage level is less than 0.05 so accept Ha
t.test(OJ_05,VC_05,var.equal=FALSE)
```
p-value is less than **alpha=0.05**, so we accept **Ha**

```{r}
# p-value at 1.0 dosage level is less than 0.05 so accept Ha
t.test(OJ_1,VC_1,var.equal=FALSE)
````
p-value is less than **alpha=0.05**, so we accept **Ha**

```{r}
# p-value at 1.0 dosage level is greater than 0.05 so fail to reject H0
t.test(OJ_2,VC_2,var.equal=FALSE)
```
p-value is greater than **alpha=0.05**, so we fail to reject **H0**

From p-values and plot, OJ seems to be more effective for dosage levels 0.5 and 1.0. But at dosage level 2.0 OJ and VC are the same.